tstrait: a quantitative trait simulator for ancestral recombination graphs

### 2023-09-16 oup-authoring-template.tex

\texttt{tstrait} offers a very efficient way of simulating quantitative traits based on an ARG. Figure \ref{fig:time} (A) shows how the simulation time scales with the number of individuals on simulated human genomes in \texttt{stdpopsim} \citep{adrion2020}. The only published code to simulate quantitative traits directly from ARGs is in \cite{martin2017}, which we compared with. Including other tools that require the entire genome matrix is not practical. For example, even with 1,000 individuals of 100 Mb human chromosome requires XX Gb of VCF. but it only takes 0.353 seconds to simulate quantitative traits with 1000 causal sites by using \texttt{tstrait}. Thus, the simulation time will be dominated by processing the VCF file. For larger samples, the problem is more

For example, the French Canadian dataset takes XX petabytes in the VCF format, but it is only XX Gb in the \texttt{tskit} compressed tree-sequence format and it only takes a few seconds to load and process by using the tree sequence data structure.

@@

>>>>

\texttt{tstrait} offers a very efficient way of simulating quantitative traits based on an ARG. Figure \ref{fig:time} (A) shows how the simulation time scales with the number of individuals on simulated human genomes in \texttt{stdpopsim} \citep{adrion2020}. The only published code to simulate quantitative traits directly from ARGs is in \cite{martin2017}, which we compared with. Including other tools that require the entire genome matrix is not practical, as the simulation time will be dominated by processing the VCF file and it would not be practically to analyze a large sample VCF file on a laptop PC. For example, the Chromosome 9 French Canadian dataset takes approximately 280 terabytes in the VCF format, but it is only 1.357 Gb in the \texttt{tskit} compressed tree-sequence format and it only takes a few minutes to load and process by using the tree sequence data structure.

@@

D.T. is supported by the Oxford Kobe Scholarship from the University of Oxford and the Euretta J. Kellett Fellowship from Columbia University.
